---
title: "hw5"
author: "Alicia Hauglie"
date: "4/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
if (file.exists("alcohol.rds")) {
 data <- readRDS("alcohol.rds")
} else {
 url <- "https://xdaiisu.github.io/ds202materials/data/ames.csv"
 data <- readr::read_csv(url)
 saveRDS(data, file="alcohol.rds")
} 
```


2. Data cleaning:
-extract geographic latitude and longitude from the variable Store Location
```{r}
library(tidyverse)
dataNew <- data %>% separate(`Store Location`, c("address1", "address2", "coords"), sep = "\n") %>% select(-address1, -address2) %>% mutate(coords = gsub("[()]", "", coords)) %>% separate(coords, c("Latitude", "Longitude"), sep = ",") 
```


-check variable types. Pick five variables that need to be converted to a different type and fix those.
```{r}
str(dataNew)
#date
library(lubridate)
dataNew$Date <- mdy(dataNew$Date)
#latitude
dataNew$Latitude <- as.numeric(dataNew$Latitude)
#longitude
dataNew$Longitude <- as.numeric(dataNew$Longitude)
#vendor number
dataNew$`Vendor Number` <- as.numeric(dataNew$`Vendor Number`)
#bottles sold
dataNew$`Bottles Sold` <- as.integer(dataNew$`Bottles Sold`)
```


-extract year, month, and day from the variable Date
```{r}
#year(dataNew$Date)
#month(dataNew$Date)
#day(dataNew$Date)
dataNew$`Day of Year` <- yday(dataNew$Date)
dataNew$Month <- month(dataNew$Date)
dataNew$Day <- day(dataNew$Date)
dataNew$Year <- year(dataNew$Date)
```




3. First overview:


-Plot a scatterplot of lat and long of store locations.
```{r}
dataNew %>% ggplot(aes(x=dataNew$Longitude, y=dataNew$Latitude))+geom_point()
```



-Provide a visual breakdown of the liquor category (by Category Name). Include volume sold in the breakdown.
```{r}
library(ggforce)
ggplot(dataNew, aes(y=`Volume Sold (Liters)`)) + geom_boxplot() + facet_wrap_paginate(~`Category Name`, ncol=5, nrow=5, scales="free_y", page = 1 )
ggplot(dataNew, aes(y=`Volume Sold (Liters)`)) + geom_boxplot() + facet_wrap_paginate(~`Category Name`, ncol=5, nrow=5, scales="free_y", page = 2 )
ggplot(dataNew, aes(y=`Volume Sold (Liters)`)) + geom_boxplot() + facet_wrap_paginate(~`Category Name`, ncol=5, nrow=5, scales="free_y", page = 3 )
ggplot(dataNew, aes(y=`Volume Sold (Liters)`)) + geom_boxplot() + facet_wrap_paginate(~`Category Name`, ncol=5, nrow=5, scales="free_y", page = 4 )
ggplot(dataNew, aes(y=`Volume Sold (Liters)`)) + geom_boxplot() + facet_wrap_paginate(~`Category Name`, ncol=5, nrow=5, scales="free_y", page = 5 )
ggplot(dataNew, aes(y=`Volume Sold (Liters)`)) + geom_boxplot() + facet_wrap_paginate(~`Category Name`, ncol=5, nrow=4, scales="free_y", page = 6 )
```

4. Find the daily sales liquor sales in Ames in 2018.
```{r}
Ames2018 <- filter(dataNew, Year == "2018")
liquor.grp.year <- group_by(Ames2018, `Day of Year`)
summarise(liquor.grp.year, sum(`Sale (Dollars)`))
```


Summarize number of sales, volume of liquor sold and amount of money spent. 
```{r}
sale.count <- group_by(dataNew, `Day of Year`)
tally(sale.count)
summarise(sale.count, sum(`Volume Sold (Liters)`), sum(`Day of Year`), sum(`Sale (Dollars)`))

```


Plot volume sold by day (use a scatterplot of volume by day and facet by month). Describe any patterns.
```{r}
library(dplyr)

ggplot(sale.count, aes(x=sale.count$Day, y=sale.count$`Volume Sold (Liters)`))+geom_point(size=.5)+facet_wrap(~Month)+xlab("Day of the Month")+ylab("Volume Sold (Liters)") 

ggplot(sale.count, aes(x=sale.count$Day, y=sale.count$`Volume Sold (Liters)`))+geom_point(size=.5)+facet_wrap(~Month)+xlab("Day of the Month")+ylab("Volume Sold (Liters)")
```


Find the dates for ISU football home games in Fall 2018. Can you see a pattern?
```{r}
#9/1/2018
#9/8/2018
#9/15/2018
#9/22/2018
#9/29/2018
#10/6/2018
#10/13/2018
#10/27/2018
#11/3/2018
#11/10/2018
#11/17/2018
#11/24/2018
#12/1/2018
```


There is a steep spike in liquor sales on game days. The day before a game has decently higher than average sales as well. We would expect this, especially in a college town. It's interesting to see the sales for home games vs away. 